---
title: "sample_095B"
output: html_document
date: "2023-06-15"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 


```{r}

#restart the R session 
#run ".rs.restartR()" in the console

# Ensure Giotto Suite is installed.
if(!"Giotto" %in% installed.packages()) {
  devtools::install_github("drieslab/Giotto@suite")
}

# Ensure GiottoData, a small, helper module for tutorials, is installed.
if(!"GiottoData" %in% installed.packages()) {
  devtools::install_github("drieslab/GiottoData")
}
# Ensure the Python environment for Giotto has been installed.
library(Giotto)
genv_exists = checkGiottoEnvironment(mini_install_path = "/projectnb/rd-spat/HOME/ivycwf/.conda/")
if(!genv_exists){
  # The following command need only be run once to install the Giotto environment.
  installGiottoEnvironment(mini_install_path= "/projectnb/rd-spat/HOME/ivycwf/.conda/")
}
```
```{r}
library(Giotto)
library(GiottoData)
# 1. set working directory
results_folder = "/projectnb/rd-spat/HOME/ivycwf/project_1/sample_095B"
my_python_path = "/projectnb/rd-spat/HOME/ivycwf/.conda/envs/giotto_env/bin/python"
instrs = createGiottoInstructions(save_dir = results_folder,
                                  save_plot = TRUE,
                                  show_plot = FALSE,
                                  python_path = my_python_path)
```


```{r Part1 : Create Giotto Object}
data_path = '/projectnb/rd-spat/HOME/ivycwf/GSE210616/095B/'
## obese upper
visium_sample_95B = createGiottoVisiumObject(
    h5_visium_path = "/projectnb/rd-spat/HOME/ivycwf/GSE210616/095B/GSM6433596_095B_filtered_feature_bc_matrix.h5",
    #expr_data = 'filtered', # Filtered data will only have in tissue
    #the image name should change to tissue_hires_image.png so it could be recongnized and functioning correctly
    h5_image_png_path = '/projectnb/rd-spat/HOME/ivycwf/GSE210616/095B/tissue_hires_image.png',
    h5_tissue_positions_path = "/projectnb/rd-spat/HOME/ivycwf/GSE210616/095B/GSM6433596_095B_tissue_positions_list.csv",
    h5_json_scalefactors_path = "/projectnb/rd-spat/HOME/ivycwf/GSE210616/095B/GSM6433596_095B_scalefactors_json.json",
    #gene_column_index = 2,
    instructions = instrs,
    do_manual_adj = F
)
```
```{r}
 visium_sample_95B <-readRDS("/projectnb/rd-spat/HOME/ivycwf/project_1/sample_095B/visium_sample_95B.rds")

```


```{r}
## show associated images with giotto object
showGiottoImageNames(visium_sample_95B) # "image" is the default name

## check metadata
pDataDT(visium_sample_95B)

## show plot
spatPlot2D(gobject = visium_sample_95B, cell_color = 'in_tissue', point_size = 2,
           show_image = T, image_name = 'image')
```

#Part 2: Process Giotto Visium Object
```{r}
## subset on spots that were covered by tissue
metadata = pDataDT(visium_sample_95B)
in_tissue_barcodes = metadata[in_tissue == 1]$cell_ID
visium_sample_95B = subsetGiotto(visium_sample_95B, cell_ids = in_tissue_barcodes)

## filter
visium_sample_95B <- filterGiotto(gobject = visium_sample_95B,
                              expression_threshold = 1,
                              feat_det_in_min_cells = 50,
                              min_det_feats_per_cell = 1000,
                              expression_values = c('raw'),
                              verbose = T)

## normalize and scale data
visium_sample_95B <- normalizeGiotto(gobject = visium_sample_95B, scalefactor = 6000, verbose = T)

## add gene & cell statistics
visium_sample_95B <- addStatistics(gobject = visium_sample_95B)

## visualize
spatPlot2D(gobject = visium_sample_95B, show_image = T, point_alpha = 0.7,
           cell_color = 'nr_feats', color_as_factor = F)
# nr_feats is in "visium_sample_95B@cell_metadata$cell$rna"
```

# Part 3: Dimention Reduction
```{r}
## highly variable features / genes (HVF)
visium_sample_95B <- calculateHVF(gobject = visium_sample_95B, save_plot = F, show_plot = T)
```


```{r}
## run PCA on expression values (default)
gene_metadata = fDataDT(visium_sample_95B) #features data
featgenes = gene_metadata[hvf == 'yes' & perc_cells > 3 & mean_expr_det > 0.4]$feat_ID

## run PCA on expression values (default)
visium_sample_95B <- runPCA(gobject = visium_sample_95B,
                       feats_to_use = featgenes)

screePlot(visium_sample_95B, ncp = 30)
```
```{r}
dimPlot2D(gobject = visium_sample_95B, dim_reduction_to_use = "pca")
```


```{r}
## run UMAP and tSNE on PCA space (default)
visium_sample_95B <- runUMAP(visium_sample_95B, dimensions_to_use = 1:10)
#plotUMAP(gobject = visium_sample_95B)
```


```{r}
visium_sample_95B <- runtSNE(visium_sample_95B, dimensions_to_use = 1:10)
#plotTSNE(gobject = visium_sample_95B)
```
# Part 4: Cluster
```{r}
## sNN network (default)
visium_sample_95B <- createNearestNetwork(gobject = visium_sample_95B, dimensions_to_use = 1:10, k = 15)

## Leiden clustering
visium_sample_95B <- doLeidenCluster(gobject = visium_sample_95B, resolution = 0.4, n_iterations = 1000)

plotUMAP(gobject = visium_sample_95B,
         cell_color = 'leiden_clus', show_NN_network = T, point_size = 2.5)
```


```{r}
# spatial and dimension plots
spatDimPlot(gobject = visium_sample_95B, cell_color = 'leiden_clus',
            dim_point_size = 2, spat_point_size = 2.5)
```


```{r}
spatDimPlot(gobject = visium_sample_95B, cell_color = 'nr_feats', color_as_factor = F,
            dim_point_size = 2, spat_point_size = 2.5)
```


```{r}
# dimension plots grouped by cluster
spatPlot2D(visium_sample_95B, cell_color = 'leiden_clus',
           coord_fix_ratio = 1)
```

# Spatial Gene expression on features for specific genes
```{r}
spatFeatPlot2D(visium_sample_95B, expression_values = "scaled",
               feats = c("APOE","FABP4","ADIPOQ","PPARG"),
               cow_n_col = 2, point_size = 1.5)
```


# Spatial Gene expression after dimension reduction for specific genes
```{r}
# umap plots
dimFeatPlot2D(visium_sample_95B, expression_values = 'scaled',
              feats = c("APOE","FABP4","ADIPOQ","PPARG"),
              cow_n_col = 2, point_size = 0.75,
              save_param = list(base_width = 10, base_height = 5))
```

#Part 5 : Run the interactive selection tool
```{r}
# could use test[,selections:=NULL] to remove the "selections" column in the data.table

## Plot spatial cells/spots # No tissue image on the background
s95B_spatPlot <- spatPlot2D(gobject = visium_sample_95B,
                          cell_color = 'leiden_clus',
                          point_size = 1.5,
                          point_alpha = 0.5,
                          show_image = FALSE)

```


```{r}
# could use test[,selections:=NULL] to remove the "selections" column in the data.table

## Plot spatial cells/spots with tissue on the background
s95B_spatPlot <- spatPlot2D(gobject = visium_sample_95B,
                          point_size = 1.5,
                          point_alpha = 0.1,
                          show_image = TRUE)
```



# Save the coordinates
```{r}
## The polygon coordinates will be saved as a data.table object
s95B_spatPlot_coordinates <- plotInteractivePolygons(s95B_spatPlot)
s95B_spatPlot_coordinates
```
```{r}
# add the polygon coordinates to the Giotto object:
## We must transform the data.table or data.frame with coordinates into a Giotto polygon object
s95B_polygons <- createGiottoPolygonsFromDfr(s95B_spatPlot_coordinates, name = 'selections_multi_regions')
s95B_polygons
```


```{r}
## Then, add the polygons to the Giotto object
visium_sample_95B <- addGiottoPolygons(gobject = visium_sample_95B,
                                  gpolygons = list(s95B_polygons))
```


```{r}
## Using the default parameters
visium_sample_95B <- addPolygonCells(visium_sample_95B,
                                polygon_name = 'selections_multi_regions')

## Let's see how it looks like now the cell_metadata
pDataDT(visium_sample_95B)

```

# Get the cells located within each polygon
```{r}
## Provide the name of the polygon to extract cells from
adipo_cells <-getCellsFromPolygon(visium_sample_95B,
                    polygon_name = 'selections_multi_regions',
                    polygons = 'multi_region_a')
adipo_cells # This is a spatial vector
              # polygon1_cells$cell_ID can get all the cells in polygon 1 

```

```{r}
## Provide the name of the polygon to extract cells from
polygon2_cells <-getCellsFromPolygon(visium_sample_95B,
                    polygon_name = 'selections',
                    polygons = 'polygon 2')
polygon2_cells # This is a spatial vector

```
```{r}
## Provide the name of the polygon to extract cells from
polygon3_cells <-getCellsFromPolygon(visium_sample_95B,
                    polygon_name = 'selections',
                    polygons = 'polygon 3')
polygon3_cells # This is a spatial vector
```
# Compare gene expression (you need to specify the genes you are interested in and if the genes are not in the vector you provided, it will give you errors)
```{r}
## provide a list of genes you want to compare
comparePolygonExpression(visium_sample_95B,
                         polygon_name = "selections_far1",
                         selected_feats = c('APOE','PPARG', 'CEBPA', 'FABP4','ADIPOQ','FASN','LPL'))

#"PPARG', 'CEBPA', 'FABP4', 'ADIPOQ', 'LEP', 'RETN', 'FASN', 'LPL', 'CFD', 'DLK1', 'CD36', 'GPD1', 'SREBF1', 'TCF21', 'ZFP423', 'CIDEC', 'CIDEA', 'PLIN1', 'FATP1', 'KLF15'"


```

# Part 6: marker gene detection for clusters
```{r}
## ------------------ ##
# Scran Markers
scran_markers_subclusters = findMarkers_one_vs_all(gobject = visium_sample_95B,
                                                   method = 'scran',
                                                   expression_values = 'normalized',
                                                   cluster_column = 'leiden_clus')
topgenes_scran = scran_markers_subclusters[, head(.SD, 2), by = 'cluster']$feats

# violinplot
violinPlot(visium_sample_95B, feats = unique(topgenes_scran), cluster_column = 'leiden_clus',
           strip_text = 8, strip_position = 'right',
           save_param = list(base_width = 5))
```
```{r}
# NO need to run this cell
# cluster heatmap
plotMetaDataHeatmap(visium_sample_95B, selected_feats = topgenes_scran,
                    metadata_cols = c('leiden_clus'))
```


```{r}
### NO need to run the cell
# umap plots
dimFeatPlot2D(visium_sample_95B, expression_values = 'scaled',
              feats = scran_markers_subclusters[, head(.SD, 1), by = 'cluster']$feats,
              cow_n_col = 3, point_size = 1,
              save_param = list(base_width = 8, base_height = 8))
```



```{r}
## Or calculate the top genes expressed within each region, then provide the result to compare polygons.
scran_results <- findMarkers_one_vs_all(visium_sample_95B,
                                        spat_unit = 'cell',
                                        feat_type = 'rna',
                                        method = 'scran',
                                        expression_values = 'normalized',
                                        cluster_column = 'selections_multi_regions',
                                        min_feats = 10)

top_genes <- scran_results[, head(.SD, 100), by = 'cluster']$feats

```
```{r}
topgenes_heatmap <- comparePolygonExpression(visium_sample_95B, polygon_name = "selections_multi_regions",
                         selected_feats = top_genes, method = "scran", expression_values = "normalized")
# This function do not save the plot
topgenes_heatmap
pdf("/projectnb/rd-spat/HOME/ivycwf/project_1/sample_095B/topgenes_heatmap_multi_regions.pdf",width = 8, height = 15)
plot(topgenes_heatmap)
dev.off()
```

```{r}
library(SeqGSEA)
#Convert symbol to ensembl ID
 adipo1_ensem <- SeqGSEA::convertSymbol2Ensembl(c("SLPI", "LYZ", "MT-CO2", "MRPL41", "LRRC26", "SLC9A3R1", "SMIM22", "TMBIM6", "MELTF", "GABRP", "ADRA2B", "MT-CO3", "COX6A1", "OAT", "AIF1L", "FBP1", "H2AFJ", "METRN", "CKS2", "MT-ND4", "RPL41", "DMAC1", "FRZB", "SOX10", "TMEM238", "TUBB4B", "CTSF", "EPCAM", "ATP1B1", "MGLL"))%>%
   dplyr::select(ensembl_gene_id)%>%
   as.vector()
 
```

```{r}
#Convert symbol to ensembl ID
 top_far_ensem <- SeqGSEA::convertSymbol2Ensembl(top30_far_genes)%>%
   dplyr::select(ensembl_gene_id)%>%
   as.vector()

```

















```{r}
# Convert the symbol ID to ensembl ID

```

# ###############################################################################
# Plot with group by: (optional)
```{r}
spatPlot2D(visium_sample_95B, cell_color = 'leiden_clus',
           group_by = 'leiden_clus', coord_fix_ratio = 1,
           cow_n_col = 6, show_legend = F,
           save_param = list(base_width = 14, base_height = 14))
#Highlight one or more groups: (optional)
spatPlot2D(visium_sample_95B, cell_color = 'leiden_clus',
           select_cell_groups = '8', coord_fix_ratio = 1, show_other_cells = TRUE,
           cell_color_code = c('8' = 'red'), other_cell_color = "grey", other_point_size = 1.5,
           save_param = list(base_width = 7, base_height = 7))
```

# Part 5: subset data (Do not need right now )
```{r}
# create and show subset
DG_subset = subsetGiottoLocs(visium_sample_95B,
                             x_max = 6500, x_min = 3000,
                             y_max = -2500, y_min = -5500,
                             return_gobject = TRUE)
# (x and y means the coordinate on the plots)

spatDimPlot(gobject = DG_subset,
            cell_color = 'leiden_clus', spat_point_size = 5)
```







# Part 7: cell-type annotation
```{r}

```






# Part 6: marker gene detection for clusters
```{r}
## ------------------ ##
## Gini markers
gini_markers_subclusters = findMarkers_one_vs_all(gobject = visium_sample_95B,
                                                  method = 'gini',
                                                  expression_values = 'normalized',
                                                  cluster_column = 'leiden_clus',
                                                  min_feats = 20,
                                                  min_expr_gini_score = 0.5,
                                                  min_det_gini_score = 0.5)
topgenes_gini = gini_markers_subclusters[, head(.SD, 2), by = 'cluster']$feats

# violinplot
violinPlot(visium_sample_95B, feats = unique(topgenes_gini), cluster_column = 'leiden_clus',
           strip_text = 6, strip_position = 'right', #strip_text = 8 is the text size
           save_param = list(base_width = 5, base_height = 10))
```


```{r}
# cluster heatmap
plotMetaDataHeatmap(visium_sample_95B, selected_feats = unique(topgenes_gini),
                    metadata_cols = c('leiden_clus'),
                    x_text_size = 10, y_text_size = 10)
```


```{r}
# umap plots
dimFeatPlot2D(visium_sample_95B, expression_values = 'scaled',
              feats = gini_markers_subclusters[, head(.SD, 1), by = 'cluster']$feats,
              cow_n_col = 4, point_size = 0.75,
              save_param = list(base_width = 8, base_height = 8))
```







```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```

