---
title: "Debug_improve_resolution"
output: html_document
date: "2023-09-22"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Check how amny clusters belongs to the edge clusters ()

```{r - **merge tiles with all clusters**}
# comb_res is a dataframe which has the tiles name and the cluster the tile belongs to
#Get all the filenames of tiles in a list 
load("/projectnb/rd-spat/HOME/ivycwf/project_1/resolution/s119B_89x103_tiles/entireHnE_kmean8_pca10_it500_star100.RData")

# Create an empty list to store the raster objects
fftiles <- list()
fftile_rows <- list()

# Function to load raster and preprocess
load_and_preprocess_raster <- function(tile_name) {
  tile <- terra::rast(tile_name)
  if (is.na(any(terra::values(tile) > 255) || any(terra::values(tile) < 0))) {
      tile_values <- terra::values(tile)
      NaN_rows <- which(apply(tile_values, 1, function(row) all(is.nan(row))))
      tile_values[NaN_rows, 1] <- 179
      tile_values[NaN_rows, 2] <- 184
      tile_values[NaN_rows, 3] <- 168
      terra::values(tile) <- tile_values
  }
  return(tile)
}

# Iterate over each kmean value and create a plot
for (kmean_val in 1:4) {
  # Get the corresponding tile names for the current kmean value
  tile_names <- lapply(comb_res[comb_res$kmeans == kmean_val, 1], function(x) paste0('/projectnb/rd-spat/HOME/ivycwf/project_1/resolution/s119B_89x103_tiles/', x))
  
  # Load each raster into the list and preprocess
  fftiles <- lapply(tile_names, load_and_preprocess_raster)
  
  # Stack the rasters in the list on top of each other
  fftiles_ls <- terra::sprc(fftiles)
  merge_image <- terra::merge(fftiles_ls)
  
  # Plot the merged raster
  plot_title <- paste("Kmeans =", kmean_val)
  terra::plot(merge_image, main = plot_title)
}

```



## There are too many clusters occupied by edge clusters, 


```{r}

```

#Build matrics caluation function
```{r}
calculate_metrics <- function(observed, predicted) {
  mse <- mean((observed - predicted)^2)
  mae <- mean(abs(observed - predicted))
  rmse <- sum((observed - predicted)^2) / sum((observed - mean(observed))^2)
  r2 <- R2(observed, predicted, form = "corr")
  pearson <- cor(predicted, observed, method = "pearson")

  cat("MAE:", mae, "\n", "MSE:", mse, "\n", 
      "RMSE:", rmse, "\n", "R-squared:", r2, "\n", "Pearson:", pearson, "\n")
}

```
#Build plot expression mapping plots
```{r}
plot_expression_map <- function(data, observed_col, predicted_col) {
  # Plot observed expression
  plot_observed <- ggplot(data, aes(x = x_cor, y = y_cor)) +
    geom_point(aes(colour = .data[[observed_col]])) +
    scale_colour_gradient2() +
    ggtitle("Observed Expression")

  # Plot predicted expression
  plot_predicted <- ggplot(data, aes(x = x_cor, y = y_cor)) +
    geom_point(aes(colour = .data[[predicted_col]])) +
    scale_colour_gradient2() +
    ggtitle("Predicted Expression")

  # Print or display the plots
  print(plot_observed)
  print(plot_predicted)
}
```

#Build train and predict model function
```{r}
model_train_predict <- function(gene_name, train_set, methods, test_set){
  control <- trainControl(method="cv",                            
                        number = 10,                            
                        summaryFunction = defaultSummary,                         
                        savePredictions = 'all')
  model <- caret::train(
                      formula(paste(gene_name, "~ .")),                      
                       data = train_set,                    
                       method = methods,                      
                       metric = "RMSE",                      
                       trControl = control) 
  
  # Make predictions on the test set
  gene_preds <- predict(model, newdata = test_set, type = "raw")
  return(list(model = model, gene_preds = gene_preds))
}

```


#Modeling
```{r}
#Model 1: LASSO model
#fit	LASSO,	need	a	random	seed	because	cross-validation	is	involved	 
#APOE_las_cv <- cv.glmnet(x = as.matrix(training_set[,-c(1, 2050, 2051)]), y = training_set$APOE, alpha = 1)
SPARC_las_cv <- cv.glmnet(x = as.matrix(training_set[,-c(1, 2050, 2051)]), y = training_set$SPARC, alpha = 1)

#Get the optimal lambda
#opt_lambda <-APOE_las_cv$lambda.min
opt_lambda <-SPARC_las_cv$lambda.min

#produce plot of test MSE by lambda value
#plot(APOE_las_cv) #why k is 1???
plot(SPARC_las_cv)

#find coefficients of best model
#APOE_las <-glmnet( x = as.matrix(training_set[,-c(1,2050, 2051)]), y = training_set$APOE, alpha = 1, lambda = opt_lambda)
SPARC_las <-glmnet( x = as.matrix(training_set[,-c(1,2050, 2051)]), y = training_set$SPARC, alpha = 1, lambda = opt_lambda)


#coef(APOE_las)
coef(SPARC_las)
#No coefficient is shown for the predictor (all original features) because lasso regression shrunk the coefficient all the way to zero. This means it was completely dropped from the model because it wasn’t influential enough.

# Make predictions on the test set
#APOE_preds <- predict(APOE_las, newx =  as.matrix(test_set[,-c(1, 2050, 2051)]), type = "response", s= opt_lambda)
SPARC_preds <- predict(SPARC_las, newx =  as.matrix(test_set[,-c(1, 2050, 2051)]), type = "response", s= opt_lambda)

calculate_metrics(cm_SPARC_lso$SPARC, cm_SPARC_lso$SPARC_pred)

plot(test_set$SPARC, SPARC_preds)
scatter.smooth(test_set$SPARC, SPARC_preds)
cor(test_set$SPARC, SPARC_preds)


```



#Model 1: #LASSO or elasticnet (glmnet) not using lm because linear regression model need more observations than variables(features) so the PC could be more reliable. -- use original features
```{r}
#Model 1 -- LASSO (another method)
model_n_preds <- model_train_predict("SPARC", training_set[,c(2:2049,2050)], "glmnet", test_set[,c(2:2049)] )
lasso_model <- model_n_preds$model
gene_preds <- model_n_preds$gene_preds
#Generate confusion matrix  
cm_gene_lso = data.frame( SPARC = test_set["SPARC"], 
                          SPARC_pred = gene_preds, 
                          tile_name = test_set[,1])
  
#Combine the original expression value and predict values with tile coordinates
cm_SPARC_lso <- merge(tile_plot_df[match(cm_gene_lso$tile_name,tile_plot_df$tile_name), c("cell_ID","tile_name","x_cor","y_cor")], 
                      cm_gene_lso, by = "tile_name")

#Calculate model evaluation matrics
calculate_metrics(cm_SPARC_lso$SPARC, cm_SPARC_lso$SPARC_pred)

#Plot predicted expression
plot_expression_map(cm_SPARC_lso, "SPARC", "SPARC_pred")

head(cm_SPARC_lso)

#Preleminary test to check the test assumptions
cor.test(cm_SPARC_lso$SPARC, cm_SPARC_lso$SPARC_pred, method=c("pearson"))
cor.test(cm_SPARC_lso$SPARC, cm_SPARC_lso$SPARC_pred, method=c("spearman"))
ggscatter(cm_SPARC_lso, x = "SPARC", y = "SPARC_pred", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          add.params = list(color = "blue", fill = "lightgray"),
          xlab = "Observed SPARC Expression", ylab = "Predicted SPARC Expression",
          size = 1.5, cor.coef.size = 5)
  #not linear relationship


```
#Test the SPARC and predicted SPARC is normally distributed
```{r}
# QQ-plot normality test for SPARC
ggqqplot(cm_SPARC_lso$SPARC, ylab = "SPARC")+
  ggtitle("Quantile-Quantile Plot of Observed SPARC Expression")

# QQ-plot normality test for predicted SPARC
ggqqplot(cm_SPARC_lso$SPARC_pred, ylab = "SPARC_pred")+
  ggtitle("Quantile-Quantile Plot of Predicted SPARC Expression")


# Shapiro-Wilk normality test for SPARC
#shapiro.test(cm_SPARC_lso$SPARC) 
# Shapiro-Wilk normality test for predicted SPARC
#shapiro.test(cm_SPARC_lso$SPARC_pred) 

ks.test(cm_SPARC_lso$SPARC, "pnorm")
ks.test(cm_SPARC_lso$SPARC_pred, "pnorm")
#Observed SPARC seems to be normally distributed
#And the predicted SPARC is normally distributed
```


#Model 1 -- LASSO for SPARC (with COL1A1 & LUM)
```{r}
#data preprocessing can also be done in train() function
SPARC_glmnet <- caret::train(SPARC~.,                      
                     data = training_set[,c(2:2049,2050:2060)],                    
                     method = "glmnet",                      
                     metric = "RMSE",                      
                     trControl = control) 

# Make predictions on the test set
SPARC_preds <- predict(SPARC_glmnet, newdata = test_set[,c(2:2049,2051:2060)], type = "raw")

#Generate confusion matrix
cm_SPARC_lso = data.frame( SPARC = test_set["SPARC"], SPARC_pred = SPARC_preds, tile_name = test_set[,1])

# Combine the original expression value and predict values with tile coordinates
cm_SPARC_lso <- merge(tile_plot_df[match(cm_SPARC_lso$tile_name, tile_plot_df$tile_name),c("cell_ID","tile_name","x_cor","y_cor")], cm_SPARC_lso, by = "tile_name")


calculate_metrics(cm_SPARC_lso$SPARC, cm_SPARC_lso$SPARC_pred)

#Plot predicted expression

plot_expression_map(cm_SPARC_lso, "SPARC", "SPARC_pred")


head(cm_SPARC_lso) 

```


# Model 1 --LASSO for COL1A1
```{r}
#Model 1 -- LASSO (another method)
gene_preds <- model_train_predict("COL1A1", training_set[,c(2:2049,2051)], "glmnet", test_set[,c(2:2049)] )
#Generate confusion matrix  
cm_gene_lso = data.frame( COL1A1 = test_set["COL1A1"], 
                          COL1A1_pred = gene_preds, 
                        tile_name = test_set[,1])

# Combine the original expression value and predict values with tile coordinates
cm_COL1A1_lso <- merge(tile_plot_df[match(cm_gene_lso$tile_name, tile_plot_df$tile_name),c("cell_ID","tile_name","x_cor","y_cor")], cm_gene_lso, by = "tile_name")


#Calculate model evaluation matrics
calculate_metrics(cm_COL1A1_lso$COL1A1, cm_COL1A1_lso$COL1A1_pred)
cor.test(cm_COL1A1_lso$COL1A1, cm_COL1A1_lso$COL1A1_pred, method=c("spearman"))
#Plot predicted expression
plot_expression_map(cm_COL1A1_lso, "COL1A1", "COL1A1_pred")

ggscatter(cm_COL1A1_lso, x = "COL1A1", y = "COL1A1_pred", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          add.params = list(color = "blue", fill = "lightgray"),
          xlab = "Observed COL1A1 Expression", ylab = "Predicted COL1A1 Expression",
          size = 1.5, cor.coef.size = 5)

```

```{r}
# normality check for COL1A1
#shapiro.test(cm_COL1A1_lso$COL1A1) 
ggqqplot(cm_COL1A1_lso$COL1A1, ylab = "COL1A1")

# normality check for predicted SPARC
#shapiro.test(cm_COL1A1_lso$COL1A1_pred) 
ggqqplot(cm_COL1A1_lso$COL1A1_pred, ylab = "COL1A1_pred")

#skewness of COL1A1 
skewness(cm_COL1A1_lso$COL1A1,na.rm=T) # is moderately skewed #not necessarily need to transformed
skewness(cm_COL1A1_lso$COL1A1_pred,na.rm=T)

library("dgof")
ks.test(cm_COL1A1_lso$COL1A1, "pnorm")
ks.test(cm_COL1A1_lso$COL1A1_pred, "pnorm")
```


# Model 1 --LASSO for LUM
```{r}
#Model 1 -- LASSO (another method)

#data preprocessing can also be done in train() function
gene_preds <- model_train_predict("LUM", training_set[,c(2:2049,2052)], "glmnet", test_set[,c(2:2049)] )
#Generate confusion matrix  
cm_gene_lso = data.frame( LUM = test_set["LUM"], 
                          LUM_pred = gene_preds, 
                        tile_name = test_set[,1])

# Combine the original expression value and predict values with tile coordinates
cm_LUM_lso <- merge(tile_plot_df[match(cm_gene_lso$tile_name, tile_plot_df$tile_name),c("cell_ID","tile_name","x_cor","y_cor")], cm_gene_lso, by = "tile_name")


#Calculate model evaluation matrics
calculate_metrics(cm_LUM_lso$LUM, cm_LUM_lso$LUM_pred)
cor.test(cm_LUM_lso$LUM, cm_LUM_lso$LUM_pred, method=c("spearman"))
#Plot predicted expression
plot_expression_map(cm_LUM_lso, "LUM", "LUM_pred")

ggscatter(cm_LUM_lso, x = "LUM", y = "LUM_pred", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          add.params = list(color = "blue", fill = "lightgray"),
          xlab = "Observed LUM Expression", ylab = "Predicted LUM Expression",
          size = 1.5, cor.coef.size = 5)

```
```{r}
# normality check for LUM
ggqqplot(cm_LUM_lso$LUM, ylab = "LUM")

# normality check for predicted SPARC
#shapiro.test(cm_COL1A1_lso$COL1A1_pred) 
ggqqplot(cm_LUM_lso$LUM_pred, ylab = "COL1A1_pred")

#skewness of COL1A1 
skewness(cm_LUM_lso$LUM,na.rm=T) # is moderately skewed #not necessarily need to transformed
skewness(cm_LUM_lso$LUM_pred,na.rm=T)
#Both normally distributed

ks.test(cm_LUM_lso$LUM, "pnorm")
ks.test(cm_LUM_lso$LUM_pred, "pnorm")

```

#Prepare df for ploting prediciton result
```{r}
#Prepare df for ploting
traing_testdf <- merge(training_set, tile_plot_df[tile_plot_df$tile_name %in% training_set$tile_name, c("tile_name","x_cor","y_cor")], by = "tile_name")
testing_testdf <- merge(test_set, tile_plot_df[tile_plot_df$tile_name %in% test_set$tile_name, c("tile_name","x_cor","y_cor")], by = "tile_name")
train_n_test_testdf <- rbind(traing_testdf, testing_testdf)

```

#Plotting results
```{r}
#Plot predict expression
ggplot(cm_SPARC_lso, aes(x=x_cor, y=y_cor)) +
  geom_point(aes(colour = SPARC_pred)) +
    scale_colour_gradient2(low = "blue", high = "red" )

#Plot visium gene expression (test set)
ggplot(cm_SPARC_lso, aes(x=x_cor, y=y_cor)) +
  geom_point(aes(colour = SPARC)) +
    scale_colour_gradient2(low = "blue", high = "red" )

#Plot visium gene expression (training set)
ggplot(traing_testdf, aes(x=x_cor, y=y_cor)) +
  geom_point(aes(colour = SPARC)) +
    scale_colour_gradient2(low = "blue", high = "red" )

#Plot visium gene expression (entire dataset)
ggplot(train_n_test_testdf, aes(x=x_cor, y=y_cor)) +
  geom_point(aes(colour = SPARC)) +
    scale_colour_gradient2(low = "blue", high = "red" )

###Check the correctness of the ploting result (training set + testing set)
combined_data <- rbind(data.frame(x_cor = cm_SPARC_lso$x_cor, y_cor = cm_SPARC_lso$y_cor, SPARC = cm_SPARC_lso$SPARC, dataset = "Lasso"),
                       data.frame(x_cor = traing_testdf$x_cor, y_cor = traing_testdf$y_cor, SPARC = traing_testdf$SPARC, dataset = "Training Set"))

# Plot combined data
ggplot(combined_data, aes(x = x_cor, y = y_cor, color = SPARC, shape = dataset)) +
  geom_point() +
  scale_colour_gradient2(low = "blue", high = "red" ) #scale_colour_gradient2(low = scales::muted("blue"), high = scales::muted("red") )

#check the distribution of variable with the histogram-- check if is's scaled
```

#Model 2: Random Forest model-- SPARC
```{r}
model_n_preds <- model_train_predict("SPARC", training_set[,c(2:2049,2050)], "rf", test_set[,c(2:2049)] )
rf_model <- model_n_preds$model
gene_preds <- model_n_preds$gene_preds

#Generate confusion matrix  
cm_gene_rf = data.frame( SPARC = test_set["SPARC"], 
                          SPARC_pred = gene_preds, 
                        tile_name = test_set[,1])
  
#Combine the original expression value and predict values with tile coordinates
cm_SPARC_rf <- merge(tile_plot_df[match(cm_gene_rf$tile_name,tile_plot_df$tile_name), c("cell_ID","tile_name","x_cor","y_cor")], 
                      cm_gene_rf, by = "tile_name")

#Calculate model evaluation matrics
calculate_metrics(cm_SPARC_rf$SPARC, cm_SPARC_rf$SPARC_pred)
cor.test(cm_SPARC_rf$SPARC, cm_SPARC_rf$SPARC_pred, method=c("spearman"))

#Plot predicted expression
plot_expression_map(cm_SPARC_rf, "SPARC", "SPARC_pred")
head(cm_SPARC_rf)

```

#Model 2: Random Forest model --COL1A1
```{r}
gene_preds <- model_train_predict("COL1A1", training_set[,c(2:2049,2051)], "rf", test_set[,c(2:2049)] )
#Generate confusion matrix  
cm_gene_rf = data.frame( COL1A1 = test_set["COL1A1"], 
                          COL1A1_pred = gene_preds, 
                        tile_name = test_set[,1])
  
#Combine the original expression value and predict values with tile coordinates
cm_COL1A1_rf <- merge(tile_plot_df[match(cm_gene_rf$tile_name,tile_plot_df$tile_name), c("cell_ID","tile_name","x_cor","y_cor")], 
                      cm_gene_rf, by = "tile_name")

#Calculate model evaluation matrics
calculate_metrics(cm_COL1A1_rf$COL1A1, cm_COL1A1_rf$COL1A1_pred)
cor.test(cm_COL1A1_rf$COL1A1, cm_COL1A1_rf$COL1A1_pred, method=c("spearman"))

#Plot predicted expression
plot_expression_map(cm_COL1A1_rf, "COL1A1", "COL1A1_pred")
head(cm_COL1A1_rf)

```

#Model 2: Random Forest model --LUM
```{r}
gene_preds <- model_train_predict("LUM", training_set[,c(2:2049,2052)], "rf", test_set[,c(2:2049)] )
#Generate confusion matrix  
cm_gene_rf = data.frame( LUM = test_set["LUM"], 
                          LUM_pred = gene_preds, 
                        tile_name = test_set[,1])
  
#Combine the original expression value and predict values with tile coordinates
cm_LUM_rf <- merge(tile_plot_df[match(cm_gene_rf$tile_name,tile_plot_df$tile_name), c("cell_ID","tile_name","x_cor","y_cor")], 
                      cm_gene_rf, by = "tile_name")

#Calculate model evaluation matrics
calculate_metrics(cm_LUM_rf$LUM, cm_LUM_rf$LUM_pred)
cor.test(cm_LUM_rf$LUM, cm_LUM_rf$LUM_pred, method=c("spearman"))

#Plot predicted expression
plot_expression_map(cm_LUM_rf, "LUM", "LUM_pred")
head(cm_LUM_rf)

```

#Model 2 : Neutral Network (neuralnet)
```{r}
library(neuralnet)

#method = "neuralnet",
#metric = "RMSE",
#trControl = control

```


#Model 3: GBM model
```{r}
model_n_preds <- model_train_predict("SPARC", training_set[,c(2:2049,2050)], "gbm", test_set[,c(2:2049)] )
gbm_model <- model_n_preds$model
gene_preds <- model_n_preds$gene_preds
#Generate confusion matrix  
cm_gene_gbm = data.frame( SPARC = test_set["SPARC"], 
                          SPARC_pred = gene_preds, 
                        tile_name = test_set[,1])
  
#Combine the original expression value and predict values with tile coordinates
cm_SPARC_gbm <- merge(tile_plot_df[match(cm_gene_gbm$tile_name,tile_plot_df$tile_name), c("cell_ID","tile_name","x_cor","y_cor")], 
                      cm_gene_gbm, by = "tile_name")

#Calculate model evaluation matrics
calculate_metrics(cm_SPARC_gbm$SPARC, cm_SPARC_gbm$SPARC_pred)
cor.test(cm_SPARC_gbm$SPARC, cm_SPARC_gbm$SPARC_pred, method=c("spearman"))
#Plot predicted expression
plot_expression_map(cm_SPARC_gbm, "SPARC", "SPARC_pred")
```

```{r}
ggscatter(cm_SPARC_gbm, x = "SPARC", y = "SPARC_pred", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          add.params = list(color = "blue", fill = "lightgray"),
          xlab = "Observed SPARC Expression", ylab = "Predicted SPARC Expression",
          size = 1.5, cor.coef.size = 5)
```


#Model 3: Random Forest GBM model (with COL1A1 and LUM)
```{r}
library(gbm) 
SPARC_gbm <- caret::train(SPARC~.,                      
                     data = training_set[,c(2:2049,2050:2060)],    #training_set[,-c(1,2007)]                
                     method = "gbm",                      
                     metric = "RSME",                      
                     trControl = control) 

# Make predictions on the test set
SPARC_gbm_preds <- predict(SPARC_gbm, newdata = test_set[,c(2:2049, 2051:2060)], type = "raw")


#Generate confusion matrix
cm_SPARC_gbm = data.frame( SPARC = test_set["SPARC"], SPARC_pred = SPARC_gbm_preds, tile_name = test_set[,1])

# Combine the original expression value and predict values with tile coordinates
cm_SPARC_gbm <- merge(tile_plot_df[match(cm_SPARC_gbm$tile_name, tile_plot_df$tile_name),c("cell_ID","tile_name","x_cor","y_cor")], cm_SPARC_gbm, by = "tile_name")

mse =  mean((cm_SPARC_gbm$SPARC - cm_SPARC_gbm$SPARC_pred)^2) 
mae = mean(abs(cm_SPARC_gbm$SPARC - cm_SPARC_gbm$SPARC_pred))
rmse = sum((cm_SPARC_gbm$SPARC - cm_SPARC_gbm$SPARC_pred)^2)/sum((cm_SPARC_gbm$SPARC - mean(cm_SPARC_gbm$SPARC))^2)
r2 = R2(cm_SPARC_gbm$SPARC, cm_SPARC_gbm$SPARC_pred, form = "corr")
cat(" MAE:", mae, "\n", "MSE:", mse, "\n", 
    "RMSE:", rmse, "\n", "R-squared:", r2)

#Polt predict expression
ggplot(cm_SPARC_gbm, aes(x=x_cor, y=y_cor)) +
  geom_point(aes(colour = SPARC_preds)) +
  scale_colour_gradient2()

ggplot(cm_SPARC_gbm, aes(x=x_cor, y=y_cor)) +
  geom_point(aes(colour = SPARC)) +
  scale_colour_gradient2()
```

#Model 3 GBM -COL1A1
```{r}
gene_preds <- model_train_predict("COL1A1", training_set[,c(2:2049,2051)], "gbm", test_set[,c(2:2049)] )
#Generate confusion matrix  
cm_gene_gbm = data.frame( COL1A1 = test_set["COL1A1"], 
                          COL1A1_pred = gene_preds, 
                        tile_name = test_set[,1])
  
#Combine the original expression value and predict values with tile coordinates
cm_COL1A1_gbm <- merge(tile_plot_df[match(cm_gene_gbm$tile_name,tile_plot_df$tile_name), c("cell_ID","tile_name","x_cor","y_cor")], 
                      cm_gene_gbm, by = "tile_name")

#Calculate model evaluation matrics
calculate_metrics(cm_COL1A1_gbm$COL1A1, cm_COL1A1_gbm$COL1A1_pred)
cor.test(cm_COL1A1_gbm$COL1A1, cm_COL1A1_gbm$COL1A1_pred, method=c("spearman"))

#Plot predicted expression
plot_expression_map(cm_COL1A1_gbm, "COL1A1", "COL1A1_pred")
```

#Model 3 GBM -LUM
```{r}
gene_preds <- model_train_predict("LUM", training_set[,c(2:2049,2052)], "gbm", test_set[,c(2:2049)] )
#Generate confusion matrix  
cm_gene_gbm = data.frame( LUM = test_set["LUM"], 
                          LUM_pred = gene_preds, 
                        tile_name = test_set[,1])
  
#Combine the original expression value and predict values with tile coordinates
cm_LUM_gbm <- merge(tile_plot_df[match(cm_gene_gbm$tile_name,tile_plot_df$tile_name), c("cell_ID","tile_name","x_cor","y_cor")], 
                      cm_gene_gbm, by = "tile_name")

#Calculate model evaluation matrics
calculate_metrics(cm_LUM_gbm$LUM, cm_LUM_gbm$LUM_pred)
cor.test(cm_LUM_gbm$LUM, cm_LUM_gbm$LUM_pred, method=c("spearman"))

#Plot predicted expression
plot_expression_map(cm_LUM_gbm, "LUM", "LUM_pred")
```

#Model4:  SVM --SPARC
```{r}
model_n_preds <- model_train_predict("SPARC", training_set[,c(2:2049,2050)], "svmPoly", test_set[,c(2:2049)] )
svm_model <- model_n_preds$model
gene_preds <- model_n_preds$gene_preds
#Generate confusion matrix  
cm_gene_svm = data.frame( SPARC = test_set["SPARC"], 
                          SPARC_pred = gene_preds, 
                        tile_name = test_set[,1])
  
#Combine the original expression value and predict values with tile coordinates
cm_SPARC_svm <- merge(tile_plot_df[match(cm_gene_svm$tile_name,tile_plot_df$tile_name), c("cell_ID","tile_name","x_cor","y_cor")], 
                      cm_gene_svm, by = "tile_name")

#Calculate model evaluation matrics
calculate_metrics(cm_SPARC_svm$SPARC, cm_SPARC_svm$SPARC_pred)

#Plot predicted expression
plot_expression_map(cm_SPARC_svm, "SPARC", "SPARC_pred")
ggscatter(cm_SPARC_svm, x = "SPARC", y = "SPARC_pred", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          add.params = list(color = "blue", fill = "lightgray"),
          xlab = "Observed SPARC Expression", ylab = "Predicted SPARC Expression",
          size = 1.5, cor.coef.size = 5)
```

#Model4:  SVM --COL1A1
```{r}
gene_preds <- model_train_predict("COL1A1", training_set[,c(2:2049,2051)], "svmPoly", test_set[,c(2:2049)] )
#Generate confusion matrix  
cm_gene_svm = data.frame( COL1A1 = test_set["COL1A1"], 
                          COL1A1_pred = gene_preds, 
                        tile_name = test_set[,1])
  
#Combine the original expression value and predict values with tile coordinates
cm_COL1A1_svm <- merge(tile_plot_df[match(cm_gene_svm$tile_name,tile_plot_df$tile_name), c("cell_ID","tile_name","x_cor","y_cor")], 
                      cm_gene_svm, by = "tile_name")

#Calculate model evaluation matrics
calculate_metrics(cm_COL1A1_svm$COL1A1, cm_COL1A1_svm$COL1A1_pred)
cor.test(cm_COL1A1_svm$COL1A1, cm_COL1A1_svm$COL1A1_pred, method=c("spearman"))

#Plot predicted expression
plot_expression_map(cm_COL1A1_svm, "COL1A1", "COL1A1_pred")
ggscatter(cm_COL1A1_svm, x = "COL1A1", y = "COL1A1_pred", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          add.params = list(color = "blue", fill = "lightgray"),
          xlab = "Observed COL1A1 Expression", ylab = "Predicted COL1A1 Expression",
          size = 1.5, cor.coef.size = 5)
```

#Model4:  SVM --LUM
```{r}
gene_preds <- model_train_predict("LUM", training_set[,c(2:2049,2052)], "svmPoly", test_set[,c(2:2049)] )
#Generate confusion matrix  
cm_gene_svm = data.frame( LUM = test_set["LUM"], 
                          LUM_pred = gene_preds, 
                        tile_name = test_set[,1])
  
#Combine the original expression value and predict values with tile coordinates
cm_LUM_svm <- merge(tile_plot_df[match(cm_gene_svm$tile_name,tile_plot_df$tile_name), c("cell_ID","tile_name","x_cor","y_cor")], 
                      cm_gene_svm, by = "tile_name")

#Calculate model evaluation matrics
calculate_metrics(cm_LUM_svm$LUM, cm_LUM_svm$LUM_pred)

#Plot predicted expression
plot_expression_map(cm_LUM_svm, "LUM", "LUM_pred")
ggscatter(cm_LUM_svm, x = "LUM", y = "LUM_pred", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          add.params = list(color = "blue", fill = "lightgray"),
          xlab = "Observed LUM Expression", ylab = "Predicted LUM Expression",
          size = 1.5, cor.coef.size = 5)

```

#Model comparisom
```{r}
model_list <- list(LASSO = lasso_model, RF = rf_model, GBM= gbm_model, SVM = svm_model) 
all_results <- resamples(model_list)
summary(all_results)
bwplot(all_results, metric = "Rsquared")

```


