---
title: "Debug_improve_resolution"
output: html_document
date: "2023-09-22"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Check how amny clusters belongs to the edge clusters ()

```{r - **merge tiles with all clusters**}
# comb_res is a dataframe which has the tiles name and the cluster the tile belongs to
#Get all the filenames of tiles in a list 
load("/projectnb/rd-spat/HOME/ivycwf/project_1/resolution/s119B_89x103_tiles/entireHnE_kmean8_pca10_it500_star100.RData")

# Create an empty list to store the raster objects
fftiles <- list()
fftile_rows <- list()

# Function to load raster and preprocess
load_and_preprocess_raster <- function(tile_name) {
  tile <- terra::rast(tile_name)
  if (is.na(any(terra::values(tile) > 255) || any(terra::values(tile) < 0))) {
      tile_values <- terra::values(tile)
      NaN_rows <- which(apply(tile_values, 1, function(row) all(is.nan(row))))
      tile_values[NaN_rows, 1] <- 179
      tile_values[NaN_rows, 2] <- 184
      tile_values[NaN_rows, 3] <- 168
      terra::values(tile) <- tile_values
  }
  return(tile)
}

# Iterate over each kmean value and create a plot
for (kmean_val in 1:4) {
  # Get the corresponding tile names for the current kmean value
  tile_names <- lapply(comb_res[comb_res$kmeans == kmean_val, 1], function(x) paste0('/projectnb/rd-spat/HOME/ivycwf/project_1/resolution/s119B_89x103_tiles/', x))
  
  # Load each raster into the list and preprocess
  fftiles <- lapply(tile_names, load_and_preprocess_raster)
  
  # Stack the rasters in the list on top of each other
  fftiles_ls <- terra::sprc(fftiles)
  merge_image <- terra::merge(fftiles_ls)
  
  # Plot the merged raster
  plot_title <- paste("Kmeans =", kmean_val)
  terra::plot(merge_image, main = plot_title)
}

```



## There are too many clusters occupied by edge clusters, 


```{r}

```




#Since there are more features than the observations, using PC would not be reliable, instead we would use original features we get from the resnet model to train the model we build
```{r}
#the matrix that combine all features of the rows (tiles) -- res_dfr 

# Convert results to matrix (image x features)
tile_names <- data.frame(tile_name = res_dfr[,1])
image_mat <- matrix(as.numeric(res_dfr[,-1]), ncol = 2048) %>% as.data.frame()
dim(image_mat) #Check how many tiles are not empty

# modify the column names avoid using number as column names
colnames(image_mat) <- paste0("f", seq_along(image_mat))

#do we have to scale or normalized the features???
#scal_image_mat <-scale(image_mat, center = T, scale = T) #sclaed by col
#scal_image_mat <-t(scale(t(image_mat), center = TRUE, scale = TRUE)) #scaled by row

#Combine the features with the corresponding tile name 
features_matrix <-cbind(tile_names, image_mat)
                                             

#features_matrix$tile_name <- gsub("/s119B_", "s119B_", features_matrix$tile_name)
input_mat <-data.frame()
input_mat <- dplyr::inner_join(features_matrix, tile_plot_df[,10:12], by = c("tile_name" = "tile_ID"))
input_mat[,1]<- sapply(input_mat$tile_name,basename) #input_mat include tile_name, 2040 features, SPARC, and MMP7 expression

#Don't need to do this every time (unless you made some changes in input_mat)
saveRDS(input_mat, file = "/projectnb/rd-spat/HOME/ivycwf/project_1/resolution/patch_tiles_4tiles/input_mat.RDS") 

```



#split the dataset 
```{r}
#load the imput_mat from RDS file
input_mat <- readRDS(file = "/projectnb/rd-spat/HOME/ivycwf/project_1/resolution/patch_tiles_4tiles/input_mat.RDS")

set.seed(123)
split = sample.split(input_mat[,1], SplitRatio = 0.8)
training_set = subset(input_mat, split == TRUE)
test_set = subset(input_mat, split == FALSE)
#sampling_index <- createDataPartition(input_mat[,1], times = 1, p = 0.8, list = FALSE) 
#training_set  <- input_mat[sampling_index, ]
#test_set <-  input_mat[-sampling_index, ]


```


#Modeling
```{r}
#Model 1: LASSO model
#fit	LASSO,	need	a	random	seed	because	cross-validation	is	involved	 
#APOE_las_cv <- cv.glmnet(x = as.matrix(training_set[,-c(1, 2050, 2051)]), y = training_set$APOE, alpha = 1)
SPARC_las_cv <- cv.glmnet(x = as.matrix(training_set[,-c(1, 2050, 2051)]), y = training_set$SPARC, alpha = 1)

#Get the optimal lambda
#opt_lambda <-APOE_las_cv$lambda.min
opt_lambda <-SPARC_las_cv$lambda.min

#produce plot of test MSE by lambda value
#plot(APOE_las_cv) #why k is 1???
plot(SPARC_las_cv)

#find coefficients of best model
#APOE_las <-glmnet( x = as.matrix(training_set[,-c(1,2050, 2051)]), y = training_set$APOE, alpha = 1, lambda = opt_lambda)
SPARC_las <-glmnet( x = as.matrix(training_set[,-c(1,2050, 2051)]), y = training_set$SPARC, alpha = 1, lambda = opt_lambda)


#coef(APOE_las)
coef(SPARC_las)
#No coefficient is shown for the predictor (all original features) because lasso regression shrunk the coefficient all the way to zero. This means it was completely dropped from the model because it wasn’t influential enough.

# Make predictions on the test set
#APOE_preds <- predict(APOE_las, newx =  as.matrix(test_set[,-c(1, 2050, 2051)]), type = "response", s= opt_lambda)
SPARC_preds <- predict(SPARC_las, newx =  as.matrix(test_set[,-c(1, 2050, 2051)]), type = "response", s= opt_lambda)

mse =  mean((test_set$SPARC - SPARC_preds)^2) 
mae = MAE(test_set$SPARC, SPARC_preds)
rmse = RMSE(test_set$SPARC, SPARC_preds)
r2 = R2(test_set$SPARC, SPARC_preds, form = "corr")
cat(" MAE:", mae, "\n", "MSE:", mse, "\n", 
    "RMSE:", rmse, "\n", "R-squared:", r2)

plot(test_set$SPARC, SPARC_preds)
scatter.smooth(test_set$SPARC, SPARC_preds)
cor(test_set$SPARC, SPARC_preds)


```

#Model 1: #LASSO or elasticnet (glmnet) not using lm because linear regression model need more observations than variables(features) so the PC could be more reliable. -- use original features
```{r}
#Model 1 -- LASSO (another method)
#Create control
control <- trainControl(method="cv",                            
                        number = 10,                            
                        summaryFunction = defaultSummary,                         
                        savePredictions = 'all')


#data preprocessing can also be done in train() function
SPARC_glmnet <- caret::train(SPARC~.,                      
                     data = training_set[,-c(1,2051)],    #training_set[,-c(1,2007)]                
                     method = "glmnet",                      
                     metric = "RMSE",                      
                     trControl = control) 

# Make predictions on the test set
SPARC_preds <- predict(SPARC_glmnet, newdata = test_set[,-c(1, 2050, 2051)], type = "raw")

#Generate confusion matrix
cm_SPARC = data.frame( SPARC = test_set[2050], SPARC_pred = SPARC_preds, tile_name = test_set[,1])

# Combine the original expression value and predict values with tile coordinates
cm_SPARC <- merge(tile_plot_df[match(cm_SPARC$tile_name, tile_plot_df$tile_name),c(3,9,13,14)], cm_SPARC, by = "tile_name")


mse =  mean((test_set$SPARC - SPARC_preds)^2) 
mae = MAE(test_set$SPARC, SPARC_preds)
rmse = RMSE(test_set$SPARC, SPARC_preds)
r2 = R2(test_set$SPARC, SPARC_preds, form = "corr")
cat(" MAE:", mae, "\n", "MSE:", mse, "\n", 
    "RMSE:", rmse, "\n", "R-squared:", r2)

```
# Model 1 --LASSO for COL5A2
```{r}
#Model 1 -- LASSO (another method)

#data preprocessing can also be done in train() function
SPARC_glmnet <- caret::train(SPARC~.,                      
                     data = training_set[,-c(1,2051)],    #training_set[,-c(1,2007)]                
                     method = "glmnet",                      
                     metric = "RMSE",                      
                     trControl = control) 

# Make predictions on the test set
SPARC_preds <- predict(SPARC_glmnet, newdata = test_set[,-c(1, 2050, 2051)], type = "raw")

#Generate confusion matrix
cm_SPARC = data.frame( SPARC = test_set[2050], SPARC_pred = SPARC_preds, tile_name = test_set[,1])

# Combine the original expression value and predict values with tile coordinates
cm_SPARC <- merge(tile_plot_df[match(cm_SPARC$tile_name, tile_plot_df$tile_name),c(3,9,13,14)], cm_SPARC, by = "tile_name")


mse =  mean((test_set$SPARC - SPARC_preds)^2) 
mae = MAE(test_set$SPARC, SPARC_preds)
rmse = RMSE(test_set$SPARC, SPARC_preds)
r2 = R2(test_set$SPARC, SPARC_preds, form = "corr")
cat(" MAE:", mae, "\n", "MSE:", mse, "\n", 
    "RMSE:", rmse, "\n", "R-squared:", r2)

```

#Prepare df for ploting prediciton result
```{r}
#Prepare df for ploting
traing_testdf <- merge(training_set, tile_plot_df[tile_plot_df$tile_name %in% training_set$tile_name, c(9,13,14)], by = "tile_name")
testing_testdf <- merge(test_set, tile_plot_df[tile_plot_df$tile_name %in% test_set$tile_name, c(9,13,14)], by = "tile_name")
train_n_test_testdf <- rbind(traing_testdf, testing_testdf)

```

#Plotting results
```{r}
#Plot predict expression
ggplot(cm_SPARC, aes(x=x_cor, y=y_cor)) +
  geom_point(aes(colour = SPARC_preds)) +
  scale_colour_gradient2()

#Polt visium gene expression (test set)
ggplot(cm_SPARC, aes(x=x_cor, y=y_cor)) +
  geom_point(aes(colour = SPARC)) +
  scale_colour_gradient2() #low = "blue", high = "red"

#Plot visium gene expression (training set)
ggplot(traing_testdf, aes(x=x_cor, y=y_cor)) +
  geom_point(aes(colour = SPARC)) +
  scale_colour_gradient2()

#Plot visium gene expression (entire dataset)
ggplot(train_n_test_testdf, aes(x=x_cor, y=y_cor)) +
  geom_point(aes(colour = SPARC)) +
  scale_colour_gradient2()

#check the distribution of varible with the histogram-- check if is's scaled
```

#Model 2: Random Forest model 
```{r}
SPARC_rf <- caret::train(SPARC~.,                      
                     data = training_set[,-c(1,2051)],    #training_set[,-c(1,2007)]                
                     method = "rf",                      
                     metric = "RMSE",                      
                     trControl = control) 

#the independent observation is violated (the data is correlated)
#SPARC_rf <- randomForest(SPARC ~ ., data = training_set[, -c(1, 2051)], importance = TRUE)

# Make predictions on the test set
SPARC_rf_preds <- predict(SPARC_rf, newdata = test_set[,-c(1, 2050, 2051)], type = "raw")

#Generate confusion matrix
cm_SPARC_rf = data.frame( SPARC = test_set[2050], SPARC_pred = SPARC_rf_preds, tile_name = test_set[,1])

# Combine the original expression value and predict values with tile coordinates
cm_SPARC_rf <- merge(tile_plot_df[match(cm_SPARC_rf$tile_name, tile_plot_df$tile_name),c(3,9,13,14)], cm_SPARC_rf, by = "tile_name")

mse =  mean((test_set$SPARC - SPARC_rf_preds)^2) 
mae = MAE(test_set$SPARC, SPARC_rf_preds)
rmse = RMSE(test_set$SPARC, SPARC_rf_preds)
r2 = R2(test_set$SPARC, SPARC_rf_preds, form = "corr") #change the R-square function
cat(" MAE:", mae, "\n", "MSE:", mse, "\n", 
    "RMSE:", rmse, "\n", "R-squared:", r2)

#Polt predict expression
ggplot(cm_SPARC_rf, aes(x=x_cor, y=y_cor)) +
  geom_point(aes(colour = SPARC_rf_preds)) +
  scale_colour_gradient2()


```
#Model 2 :Stacked AutoEncoder Deep Neutral Network (dnn) #all the predict values are the same
```{r}
library(deepnet)
# Assuming 'training_set' and 'test_set' are your data

SPARC_rf <- caret::train(SPARC~.,                      
                     data = training_set[,-c(1,2051)],    #training_set[,-c(1,2007)]  
                     method = "dnn",                      
                     metric = "RMSE",                      
                     trControl = control) 

# Make predictions on the test set
SPARC_dnn_preds <- predict(SPARC_rf, newdata = test_set[,-c(1, 2050, 2051)], type = "raw")

#Generate confusion matrix
cm_SPARC_dnn = data.frame( SPARC = test_set[2050], SPARC_pred = SPARC_dnn_preds, tile_name = test_set[,1])

# Combine the original expression value and predict values with tile coordinates
cm_SPARC_dnn <- merge(tile_plot_df[match(cm_SPARC_dnn$tile_name, tile_plot_df$tile_name),c(3,9,13,14)], cm_SPARC_dnn, by = "tile_name")

mse =  mean((test_set$SPARC - SPARC_dnn_preds)^2) 
mae = MAE(test_set$SPARC, SPARC_dnn_preds)
rmse = RMSE(test_set$SPARC, SPARC_dnn_preds)
r2 = R2(test_set$SPARC, SPARC_dnn_preds, form = "corr") #change the R-square function
cat(" MAE:", mae, "\n", "MSE:", mse, "\n", 
    "RMSE:", rmse, "\n", "R-squared:", r2)

#Polt predict expression
ggplot(cm_SPARC_dnn, aes(x=x_cor, y=y_cor)) +
  geom_point(aes(colour = SPARC_dnn_preds)) +
  scale_colour_gradient2()



```



#Model 2 : Neutral Network (neuralnet)
```{r}
library(neuralnet)

SPARC_nn <- caret::train(
  SPARC ~ .,
  data = training_set[,-c(1,2051)],
  method = "neuralnet",
  metric = "RMSE",
  trControl = control
)

SPARC_nn_preds <-predict(SPARC_nn, newdata = test_set[,-c(2050)])

#Generate confusion matrix
cm_SPARC_nn = data.frame( SPARC = test_set[2050], SPARC_pred = SPARC_nn_preds, tile_name = test_set[,1])

#Generate confusion matrix
cm_SPARC_nn = data.frame( SPARC = test_set[2050], SPARC_pred = SPARC_nn_preds, tile_name = test_set[,1])

cm_SPARC_nn <- merge(tile_plot_df[match(cm_SPARC_nn$tile_name, tile_plot_df$tile_name),c(3,9,13,14)], cm_SPARC_nn, by = "tile_name")

mse =  mean((test_set$SPARC - SPARC_nn_preds)^2) 
mae = MAE(test_set$SPARC, SPARC_nn_preds)
rmse = RMSE(test_set$SPARC, SPARC_nn_preds)
r2 = R2(test_set$SPARC, SPARC_nn_preds, form = "corr")
cat(" MAE:", mae, "\n", "MSE:", mse, "\n", 
    "RMSE:", rmse, "\n", "R-squared:", r2)
head(cm_SPARC_nn)
```



#Model 3: Random Forest GBM model
```{r}
library(gbm) 
SPARC_gbm <- caret::train(SPARC~.,                      
                     data = training_set[,-c(1,2051)],    #training_set[,-c(1,2007)]                
                     method = "gbm",                      
                     metric = "RSME",                      
                     trControl = control) 

# Make predictions on the test set
SPARC_gbm_preds <- predict(SPARC_gbm, newdata = test_set[,-c(1, 2050, 2051)], type = "raw")


#Generate confusion matrix
cm_SPARC_gbm = data.frame( SPARC = test_set[2050], SPARC_pred = SPARC_gbm_preds, tile_name = test_set[,1])

# Combine the original expression value and predict values with tile coordinates
cm_SPARC_gbm <- merge(tile_plot_df[match(cm_SPARC_gbm$tile_name, tile_plot_df$tile_name),c(3,9,13,14)], cm_SPARC_gbm, by = "tile_name")

mse =  mean((test_set$SPARC - SPARC_gbm_preds)^2) 
mae = MAE(test_set$SPARC, SPARC_gbm_preds)
rmse = RMSE(test_set$SPARC, SPARC_gbm_preds)
r2 = R2(test_set$SPARC, SPARC_gbm_preds, form = "corr")
cat(" MAE:", mae, "\n", "MSE:", mse, "\n", 
    "RMSE:", rmse, "\n", "R-squared:", r2)

#Polt predict expression
ggplot(cm_SPARC_gbm, aes(x=x_cor, y=y_cor)) +
  geom_point(aes(colour = SPARC_gbm_preds)) +
  scale_colour_gradient2()
```


#Model4:  SVM
```{r}
#Use the SVM model 
SPARC_svm <- caret::train(SPARC~.,                      
                     data = training_set[,-c(1,2051)],    #training_set[,-c(1,2007)]
                     method = "svmPoly",                      
                     metric = "RMSE",                      
                     trControl = control) 

# Predicting the Test set results (without giving it real expression value)
#APOE_pred = predict(svm_mod_APOE , newdata = test_set[,-c(1,10,11)])
SPARC_svm_preds = predict(SPARC_svm , newdata = test_set[,-c(1,2050, 2051)])

#Generate confusion matrix
cm_SPARC_svm = data.frame( SPARC = test_set[2050], SPARC_pred = SPARC_svm_preds, tile_name = test_set[,1]) #APOE = test_set[10]


# Combine the original expression value and predict values with tile coordinates
cm_SPARC_svm <- merge(tile_plot_df[match(cm_SPARC_svm$tile_name, tile_plot_df$tile_name),c(3,9,13,14)], cm_SPARC_svm, by = "tile_name")


mse =  mean((test_set$SPARC - SPARC_svm_preds)^2) 
mae = MAE(test_set$SPARC, SPARC_svm_preds)
rmse = RMSE(test_set$SPARC, SPARC_svm_preds)
r2 = R2(test_set$SPARC, SPARC_svm_preds, form = "corr")
cat(" MAE:", mae, "\n", "MSE:", mse, "\n", 
    "RMSE:", rmse, "\n", "R-squared:", r2)

#Polt predict expression
ggplot(cm_SPARC_svm, aes(x=x_cor, y=y_cor)) +
  geom_point(aes(colour = SPARC_svm_preds)) +
  scale_colour_gradient2()
```
#Using the following code to debug 
```{r}

plot(test_set$APOE, APOE_rf_preds)
scatter.smooth(test_set$APOE, APOE_rf_preds)
cor(test_set$APOE, APOE_rf_preds)
```


