---
title: "sample_119B"
output:
  pdf_document: default
  rmarkdown::github_document
date: "2023-06-16"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
# Ensure Giotto Suite is installed.
if(!"Giotto" %in% installed.packages()) {
  devtools::install_github("drieslab/Giotto@suite")
}

# Ensure GiottoData, a small, helper module for tutorials, is installed.
if(!"GiottoData" %in% installed.packages()) {
  devtools::install_github("drieslab/GiottoData")
}
# Ensure the Python environment for Giotto has been installed.
library(Giotto)
genv_exists = checkGiottoEnvironment(mini_install_path = "/projectnb/rd-spat/HOME/ivycwf/.conda/")
if(!genv_exists){
  # The following command need only be run once to install the Giotto environment.
  installGiottoEnvironment(mini_install_path= "/projectnb/rd-spat/HOME/ivycwf/.conda/")
}

```


```{r}
library(Giotto)
library(GiottoData)
# 1. set working directory
results_folder = "/projectnb/rd-spat/HOME/ivycwf/project_1/sample_119B"
my_python_path = "/projectnb/rd-spat/HOME/ivycwf/.conda/envs/giotto_env/bin/python"
instrs = createGiottoInstructions(save_dir = results_folder,
                                  save_plot = TRUE,
                                  show_plot = FALSE,
                                  python_path = my_python_path)

```

```{r Part1 : Create Giotto Object}
## obese upper
visium_sample_119B = createGiottoVisiumObject(
    h5_visium_path = "/projectnb/rd-spat/HOME/ivycwf/GSE210616/119B/GSM6433605_119B_filtered_feature_bc_matrix.h5",
    #the image name should change to tissue_hires_image.png so it could be recongnized and functioning correctly
    h5_image_png_path = '/projectnb/rd-spat/HOME/ivycwf/GSE210616/119B/tissue_hires_image.png',
    h5_tissue_positions_path = "/projectnb/rd-spat/HOME/ivycwf/GSE210616/119B/GSM6433605_119B_tissue_positions_list.csv",
    h5_json_scalefactors_path = "/projectnb/rd-spat/HOME/ivycwf/GSE210616/119B/GSM6433605_119B_scalefactors_json.json",
    instructions = instrs,
    do_manual_adj = F
)
```

```{r}
## show associated images with giotto object
showGiottoImageNames(visium_sample_119B) # "image" is the default name

## check metadata
pDataDT(visium_sample_119B)

## show plot
spatPlot2D(gobject = visium_sample_119B, cell_color = 'in_tissue', point_size = 2,
           show_image = T, image_name = 'image')
```
#Part 2: Process Giotto Visium Object
```{r}
## subset on spots that were covered by tissue
metadata = pDataDT(visium_sample_119B)
in_tissue_barcodes = metadata[in_tissue == 1]$cell_ID
visium_sample_119B = subsetGiotto(visium_sample_119B, cell_ids = in_tissue_barcodes)

## filter
visium_sample_119B <- filterGiotto(gobject = visium_sample_119B,
                              expression_threshold = 1, # 2 or 3 
                              feat_det_in_min_cells = 50,
                              min_det_feats_per_cell = 1000,#1100
                              expression_values = c('raw'),
                              verbose = T)

## normalize
visium_sample_119B <- normalizeGiotto(gobject = visium_sample_119B, scalefactor = 6000, verbose = T)

## add gene & cell statistics
visium_sample_119B <- addStatistics(gobject = visium_sample_119B)

## visualize
spatPlot2D(gobject = visium_sample_119B, show_image = T, point_alpha = 0.7,
           cell_color = 'nr_feats', color_as_factor = F, axis_text = 12, axis_title = 12)
# nr_feats is in "visium_sample_95B@cell_metadata$cell$rna"

filt_3_1100_spots <- pDataDT(visium_sample_119B)
#save(visium_sample_119B, filt_3_1100_spots, file = "/projectnb/rd-spat/HOME/ivycwf/project_1/increase_res/patches/visium_119B_spots.RData")
```


# Part 3: Dimention Reduction
```{r}
## highly variable features / genes (HVF)
visium_sample_119B <- calculateHVF(gobject = visium_sample_119B, save_plot = F, show_plot = T) #save_plot = T
```

```{r}
## run PCA on expression values (default)
gene_metadata = fDataDT(visium_sample_119B)
featgenes = gene_metadata[hvf == 'yes' & perc_cells > 3 & mean_expr_det > 0.4]$feat_ID

## run PCA on expression values (default)
visium_sample_119B <- runPCA(gobject = visium_sample_119B,
                       feats_to_use = featgenes)

screePlot(visium_sample_119B, ncp = 30)

dimPlot2D(gobject = visium_sample_119B, dim_reduction_to_use = "pca",point_size = 2, axis_text = 12, axis_title = 12, save_plot = F)
```


```{r}
## run UMAP and tSNE on PCA space (default)
visium_sample_119B <- runUMAP(visium_sample_119B, dimensions_to_use = 1:10)
plotUMAP(gobject = visium_sample_119B, point_size = 2, axis_text = 12, axis_title = 12)
```


```{r}
visium_sample_119B <- runtSNE(visium_sample_119B, dimensions_to_use = 1:10)
plotTSNE(gobject = visium_sample_119B)
```

# Part 4: Cluster
```{r}
## sNN network (default)
visium_sample_119B <- createNearestNetwork(gobject = visium_sample_119B, dimensions_to_use = 1:10, k = 15)

## Leiden clustering
visium_sample_119B <- doLeidenCluster(gobject = visium_sample_119B, resolution = 0.5, n_iterations = 1000)

plotUMAP(gobject = visium_sample_119B,
         cell_color = 'leiden_clus', show_NN_network = T, point_size = 2.5 , axis_text = 12, axis_title = 12)
```

```{r}
# spatial and dimension plots
spatDimPlot(gobject = visium_sample_119B, cell_color = 'leiden_clus',
            dim_point_size = 2, spat_point_size = 2.5)
```


```{r}
spatDimPlot(gobject = visium_sample_119B, cell_color = 'nr_feats', color_as_factor = F,
            dim_point_size = 2, spat_point_size = 2.5)
```


```{r}
# dimension plots grouped by cluster
spatPlot2D(visium_sample_119B, cell_color = 'leiden_clus', show_image = T, save_plot = F,
           coord_fix_ratio = 1, point_size = 2, axis_text = 12, axis_title = 12)
```
# Spatial Gene expression on features for specific genes
```{r}
#Shared genes related to adipose tissue
spatFeatPlot2D(visium_sample_119B, expression_values = "scaled",
               feats = c("APOE","TMEM120B","HOXC9","EPSTI1","MTUS1"),
               cow_n_col = 2, point_size = 1.5)
```


```{r}
#WAT
spatFeatPlot2D(visium_sample_119B, expression_values = "scaled",
               feats = c("FABP4","Fbxo31","LPL","MPZL2", "NR1H3","RB1","RBL1"),
               cow_n_col = 2, point_size = 1.5)

```


```{r}
#Beige
spatFeatPlot2D(visium_sample_119B, expression_values = "scaled",
               feats = c("TBX1","SP100","SLC27A1","NR2F6","CD40"),
               cow_n_col = 2, point_size = 1.5)

```


```{r}
#BAT
spatFeatPlot2D(visium_sample_119B, expression_values = "scaled",
               feats = c("BMP7","PREX1","EVA1A","EVA1B","EVA1C"),
               cow_n_col = 2, point_size = 1.5)

```



# Spatial Gene expression after dimension reduction for specific genes
```{r}
# umap plots
dimFeatPlot2D(visium_sample_119B, expression_values = 'scaled',
              feats = c("APOE","FABP4"),
              cow_n_col = 2, point_size = 1.5,
              save_param = list(base_width = 10, base_height = 5))

```

#Testing more gene, especially more beige or brown fat cell marker genes


# Do the Spatial co-expression module
# Calculate spatially variable genes
```{r}
#Create a spatial network first
#Create a spatial network based on cell centroid physical distances.
visium_sample_119B <- createSpatialNetwork(gobject = visium_sample_119B,
                                     method = 'kNN', k = 15, #5
                                     maximum_distance_knn = 400,
                                     name = 'spatial_network')

showGiottoSpatNetworks(visium_sample_119B)

spatPlot2D(gobject = visium_sample_119B,  show_network= T,
           network_color = 'blue', spatial_network_name = 'spatial_network', axis_text = 12, axis_title = 12)

```


#Identify top ranks spatially correlated genes
```{r}
#you can identify spatial genes with any type of method, here we will use the binSpect method
## rank binarization

#BinSpect (Binary Spatial Extraction of genes) is a fast computational method that identifies genes with a spatially coherent expression pattern.
ranktest = binSpect(visium_sample_119B, bin_method = 'rank',
                    calc_hub = T, hub_min_int = 5,
                    spatial_network_name = 'spatial_network')

#calc_hub -> calculate the number of hub cells
#hub_min_int -> minimum number of cell-cell interactions for a hub cell

spatFeatPlot2D(visium_sample_119B, expression_values = 'scaled',
               feats = ranktest$feats[1:9], cow_n_col = 3, point_size = 1.5 ,legend_text = 6, axis_text = 6)


```

```{r}
#Top 50 rank spatial genes
ranktest$feats[1:50]
```

#Identify top spatially correlated genes to the specific gene of interest
```{r}
# 1. cluster the top 700 spatial genes into 20 clusters
ext_spatial_genes = ranktest[1:700,]$feats

# here we use existing detectSpatialCorGenes function to calculate pairwise distances between genes (but set network_smoothing=0 to use default clustering)
#detectSpatialCorFeats() -- Detect features that are spatially correlated
spat_cor_netw_DT = detectSpatialCorFeats(visium_sample_119B,
                                         method = 'network',
                                         spatial_network_name = 'spatial_network',
                                         subset_feats = ext_spatial_genes)

# 2. identify most similar spatially correlated genes for one gene #Shows and filters spatially correlated features
#Find the top n genes that is most spatially related to the gene of your interest
top10_APOE_genes = showSpatialCorFeats(spat_cor_netw_DT, feats = 'APOE', show_top_feats = 10)
top10_APOE_genes                                                                    #show top related features per gene

```


```{r}
# plot the spatial expression of top n genes that is most spatially related to the gene of your interest

#pdf("/projectnb/rd-spat/HOME/ivycwf/project_1/sample_119B/APOE_spat_related_genes.pdf",width = 8, height = 15)

spatFeatPlot2D(visium_sample_119B, expression_values = 'scaled',
               feats = top10_APOE_genes$variable[1:9], point_size = 1.8, axis_text = 7, axis_title = 9, cow_rel_h = 7 , cow_rel_w = 9 , legend_text = 7,cow_n_col = 3, save_plot = F)

#dev.off()
```

#Use heatmap to confirm that these 10 spatially correlated genes with AOPE gene
```{r}
## provide a list of APOE spatially correlated genes you want to compare
comparePolygonExpression(visium_sample_119B,
                         polygon_name = "selections_multi_regions",
                         selected_feats = c('APOE','APOC1','TYROBP','C1QB','C1QA','IFI27','C1QC','CD74','TREM2','CD52'))
```

```{r} 
#FABP4 related spatial gene

top10_FABP4_genes = showSpatialCorFeats(spat_cor_netw_DT, feats = 'FABP4', show_top_feats = 10)
top10_FABP4_genes  
 
spatFeatPlot2D(visium_sample_119B, expression_values = 'scaled',
               feats = top10_FABP4_genes$variable[1:10], point_size = 2, axis_text =9, axis_title = 10, cow_rel_h = 7,
               cow_rel_w = 9 , legend_text = 7,cow_n_col = 3, save_plot = F)

```


# Cluster and visualize spatially correlated genes
```{r}
# use the heatmap to identify spatial co-expression modules and adjust the number of clusters (k) accordingly
# 3. identify potenial spatial co-expression (from top n genes you subset before)
# cluster spatial genes
# clusterSpatialCorFeats() -- Cluster based on spatially correlated features # k = number of clusters to extract
spat_cor_netw_DT = clusterSpatialCorFeats(spat_cor_netw_DT, name = 'spat_netw_clus', k =9) #7

# The following can get the gene names and the # of the cluster those gene belongs to!
#spat_cor_netw_DT$cor_clusters$spat_netw_clus[spat_cor_netw_DT$cor_clusters$spat_netw_clus == 7]

# If you only want the gene names 
# names(spat_cor_netw_DT$cor_clusters$spat_netw_clus[spat_cor_netw_DT$cor_clusters$spat_netw_clus == 7])


# visualize clusters
heatmSpatialCorFeats(visium_sample_119B,
                     spatCorObject = spat_cor_netw_DT,
                     use_clus_name = 'spat_netw_clus',
                     heatmap_legend_param = list(title = NULL),
                     save_param = list(base_height = 6, base_width = 8, units = 'cm'))
```


```{r}
# 4. rank spatial correlated clusters and show genes for selected clusters
# Rank spatial correlated clusters according to correlation structure
#Generate the plot that represent ......
netw_ranks = rankSpatialCorGroups(visium_sample_119B,
                                  spatCorObject = spat_cor_netw_DT, use_clus_name = 'spat_netw_clus',
                                  save_param = list(  base_height = 3, base_width = 5), show_plot = T)
netw_ranks
```


```{r}
# Shows and filters spatially correlated features

# selected_clusters -> subset of clusters to show # show_top_feats -> show top features per gene #selected_clusters -> the value means the specific cluster you would like to know
top_netw_spat_cluster = showSpatialCorFeats(spat_cor_netw_DT, use_clus_name = 'spat_netw_clus', selected_clusters = 4, show_top_feats = 2)
top_netw_spat_cluster
```

#Create metagenes/co-expression modules and visualize
```{r}
#extract a balanced set of genes from each co-expression module. #aggregate genes into metagenes. 
cluster_genes = getBalancedSpatCoexpressionFeats(spat_cor_netw_DT, maximum = 50)
                                                # maximum represent the max number of genes extract from each cluster
cluster_genes
```


```{r}
visium_sample_119B = createMetafeats(visium_sample_119B, feat_clusters = cluster_genes, name = 'cluster_metagene')
# the Metafeats info will be store in the "visium_sample_119B@spatial_enrichment$cell$rna$cluster_metagene@enrichDT" -- normalized rna expression values were used in this function

# Visualize cells according to spatial coordinates
spatCellPlot2D(visium_sample_119B,
             spat_enr_names = 'cluster_metagene',
             cell_annotation_values = as.character(c(1:9)), #	numeric cell annotation columns  #1:7
             cow_n_col = 3, point_size = 1.5 ,legend_text = 8, axis_text = 8 , axis_title = 10)

```


# Subset the genes in the clusters that we are interested in 
```{r}
#Subet cluster 3 & 4 that match to the adipose tissue # ext_spatial_genes = ranktest[1:500,]$feats
metagenes_3 <-names(cluster_genes[cluster_genes == 3])
metagenes_4 <-names(cluster_genes[cluster_genes == 4])
metagenes_6 <-names(cluster_genes[cluster_genes == 6])

```
# Convert symbol to ensembl ID
```{r}
library(SeqGSEA)
cluser3_ensem <- SeqGSEA::convertSymbol2Ensembl(metagenes_3)
cluser4_ensem <- SeqGSEA::convertSymbol2Ensembl(metagenes_4)
cluser6_ensem <- SeqGSEA::convertSymbol2Ensembl(metagenes_6)
cluser3_ensem
cluser4_ensem
cluser6_ensem
```


```{r}
### This is only selecting 40 features in each cluster
library(SeqGSEA)
#Convert symbol to ensembl ID
 cluster3_ensem <- SeqGSEA::convertSymbol2Ensembl(c("IGFBP7", "SULF1", "TAGLN", "SFRP2", "CD74", "C1QB", "HTRA3", "SPARC", "MYL6", "COL4A2", "COL4A1", "C1QA", "COL1A1", "C1QC", "THY1", "IFI27", "ACTB", "VCAN", "ACTA2", "MMP11", "APOE", "CST3", "IFI30", "TREM2", "TYROBP", "HTRA1", "COL1A2", "TPM1", "APOC1", "SLCO2B1", "SUGCT", "CD14", "ASPN", "PLVAP", "COL5A1", "PECAM1", "MFAP2", "COMP", "TPSB2", "SCT"))%>%
   dplyr::select(ensembl_gene_id)%>%
   as.vector()
 
  cluster4_ensem <- SeqGSEA::convertSymbol2Ensembl(c("TMSB4X", "S100A1", "CRABP2", "KRT23", "TUBB3", "KRT7", "TUBB2A", "DMKN", "LSR", "MMP7", "SELENOW", "ACTG2", "S100A11", "CRYAB", "DNAJC15", "S100A16", "RHOC", "TMED9", "CALM2", "ELN", "COTL1", "PFN1", "TAGLN2", "CFL1", "TUBA1A", "TNFRSF12A", "HSPB1", "UCHL1", "PYCARD", "SERF2", "CLDN4", "CHI3L2", "RBP1", "DSG1", "DBNDD1", "SDC1", "S100A6", "TMSB10", "S100A13", "PPIA"))%>%
   dplyr::select(ensembl_gene_id)%>%
   as.vector()

```


```{r}
#Subet cluster 3 & 7 that match to the adipose tissue # ext_spatial_genes = ranktest[1:700,]$feats
metagenes_3 <-names(cluster_genes[cluster_genes == 3])
metagenes_7 <-names(cluster_genes[cluster_genes == 7])
metagenes_6 <-names(cluster_genes[cluster_genes == 6])
metagenes_1 <-names(cluster_genes[cluster_genes == 1])
```

```{r}
# Convert symbol to ensembl ID
library(SeqGSEA)
cluser1_ensem <- SeqGSEA::convertSymbol2Ensembl(metagenes_1)
cluser3_ensem <- SeqGSEA::convertSymbol2Ensembl(metagenes_3)
cluser7_ensem <- SeqGSEA::convertSymbol2Ensembl(metagenes_7)
cluser6_ensem <- SeqGSEA::convertSymbol2Ensembl(metagenes_6)
cluser3_ensem
cluser7_ensem
cluser6_ensem
cluser1_ensem
```

```{r}
### This is only selecting 40 features in each cluster
cluser3_ensem$ensembl_gene_id
cluser7_ensem$ensembl_gene_id
cluser6_ensem$ensembl_gene_id
cluser1_ensem$ensembl_gene_id
```




# Spatially informed clustering
```{r}
#Here we illustrate how to use information from #5 Spatial correlated genes# as input for clustering using the standard pipeline (PCA > nearest-neighbors > leiden clusters).
all_spatial_genes = names(cluster_genes)

visium_sample_119B <- runPCA(gobject = visium_sample_119B,
                      feats_to_use = all_spatial_genes,
                      name = 'meta_clusters_pca')
```


```{r}
visium_sample_119B <- runUMAP(visium_sample_119B,
                       dim_reduction_name = 'meta_clusters_pca',
                       dimensions_to_use = 1:10, # 1:20
                       name = 'meta_clusters_umap',verbose = T)
screePlot(visium_sample_119B, ncp = 30)
```


```{r}
visium_sample_119B <- createNearestNetwork(gobject = visium_sample_119B,
                                    dim_reduction_name = 'meta_clusters_pca',
                                    dimensions_to_use = 1:10, k = 5, #1:20
                                    name = 'meta_clusters_NN',verbose = T)
```


```{r}
visium_sample_119B <- doLeidenCluster(gobject = visium_sample_119B,
                               network_name = 'meta_clusters_NN',
                               resolution = 0.15, n_iterations = 1000, #resolution = 0.15
                               name = 'meta_clusters_leiden')
```

```{r}
spatPlot2D(visium_sample_119B,
           cell_color = 'meta_clusters_leiden', point_size = 3, show_image = T,
           legend_text = 10, axis_text = 10 , axis_title = 11)
```

```{r}
plotUMAP(gobject = visium_sample_119B,
         cell_color = 'meta_clusters_leiden', show_NN_network = T, point_size = 3)
```

```{r}
saveRDS(visium_sample_119B, file = "/projectnb/rd-spat/HOME/ivycwf/project_1/sample_119B/visium_sample_119B.rds")
```


```{r}
visium_sample_119B<- readRDS("/projectnb/rd-spat/HOME/ivycwf/project_1/sample_119B/visium_sample_119B.rds")

```

#######################################################################################
#Part 5 : Run the interactive selection tool
```{r}
# could use test[,selections:=NULL] to remove the "selections" column in the data.table

## Plot spatial cells/spots # No tissue image on the background
spatPlot2D(gobject = visium_sample_119B,
                          cell_color = 'leiden_clus',
                          point_size = 1.5,
                          point_alpha = 0.5,
                          show_image = FALSE)

```


```{r}
# could use test[,selections:=NULL] to remove the "selections" column in the data.table

## Plot spatial cells/spots with tissue on the background
## Prepare for the ploygon selection
s119B_spatPlot <- spatPlot2D(gobject = visium_sample_119B,
                      point_size = 2,
                      point_alpha = 0.5,
                      show_image = TRUE)

s119B_spatPlot
```


# Select ploygon
# Save the coordinates
```{r}
## The polygon coordinates will be saved as a data.table object
s119B_spatPlot_coordinates <- plotInteractivePolygons(s119B_spatPlot)
s119B_spatPlot_coordinates
```
```{r}
# add the polygon coordinates to the Giotto object:
## We must transform the data.table or data.frame with coordinates into a Giotto polygon object
s119B_polygons <- createGiottoPolygonsFromDfr(s119B_spatPlot_coordinates, name = 'selections_multi_regions')
s119B_polygons
```


```{r}
## Then, add the polygons to the Giotto object
visium_sample_119B <- addGiottoPolygons(gobject = visium_sample_119B,
                                  gpolygons = list(s119B_polygons))
```


```{r}
## Using the default parameters
visium_sample_119B <- addPolygonCells(visium_sample_119B,
                                polygon_name = 'selections_multi_regions')

## Let's see how it looks like now the cell_metadata
pDataDT(visium_sample_119B)

```
# Get the cells located within each polygon
```{r}
## Provide the name of the polygon to extract cells from
adipo_cells <-getCellsFromPolygon(visium_sample_119B,
                    polygon_name = 'selections_multi_regions',
                    polygons = 'adipo')
adipo_cells # This is a spatial vector
            # polygon1_cells$cell_ID can get all the cells in adipo polygon

```

```{r}
## Provide the name of the polygon to extract cells from
adjacent_cells <-getCellsFromPolygon(visium_sample_119B,
                    polygon_name = 'selections_multi_regions',
                    polygons = 'adjacent')
adjacent_cells # This is a spatial vector

```
```{r}
## Provide the name of the polygon to extract cells from
far_cells <-getCellsFromPolygon(visium_sample_119B,
                    polygon_name = 'selections_multi_regions',
                    polygons = 'far')
far_cells # This is a spatial vector
```
# Compare gene expression (you need to specify the genes you are interested in and if the genes are not in the vector you provided, it will give you errors)
```{r}
## provide a list of genes you want to compare
comparePolygonExpression(visium_sample_119B,
                         polygon_name = "selections_multi_regions",
                         selected_feats = c('APOE','FABP4','CEBPA','FASN','LPL','GPD1', 'SREBF1'))

#s119_feats <-visium_sample_119B@feat_ID$rna
#symbols_to_check %in% s119_feats 


```


```{r}
## provide a list of genes that are highly express across all ploygons of top30 genes in each polygon 
comparePolygonExpression(visium_sample_119B,
                         polygon_name = "selections_multi_regions",
                         selected_feats = c('CRABP2', 'MMP7', 'TMSB4X', 'TAGLN2', 'SELENOW', 'S100A16', 'S100A11', 'S100A13','S100A1', 'TMEM176A','KRT8','CLU','APOE', 'FABP4'))

#s119_feats <-visium_sample_119B@feat_ID$rna
#symbols_to_check %in% s119_feats 
```
# Part 6: marker gene detection for clusters

```{r}
## Or calculate the top genes expressed within each region, then provide the result to compare polygons.
scran_results <- findMarkers_one_vs_all(visium_sample_119B,
                                        spat_unit = 'cell',
                                        feat_type = 'rna',
                                        method = 'scran',
                                        expression_values = 'normalized',
                                        cluster_column = 'selections_multi_regions',
                                        min_feats = 10)

top_genes <- scran_results[, head(.SD, 100), by = 'cluster']$feats
top30_genes <- scran_results[, head(.SD, 30), by = 'cluster']$feats
```

```
##top_genes
  [1] "CRABP2"       "CFL1"         "S100A16"      "MMP7"         "SFRP2"        "COTL1"       
  [7] "SDC1"         "TMSB4X"       "TUBA1A"       "ACTB"         "DMKN"         "SPARC"       
 [13] "SULF1"        "TMED9"        "TPM1"         "TMSB10"       "TAGLN"        "MYL6"        
 [19] "IGFBP7"       "ACTG2"        "ID3"          "PFN1"         "TAGLN2"       "CCND1"       
 [25] "KRT5"         "SELENOW"      "SERF2"        "OST4"         "CAVIN3"       "COLGALT1"    
 [31] "ACTA2"        "KRT23"        "HTRA1"        "COL3A1"       "TPM4"         "HTRA3"       
 [37] "S100A11"      "COL5A2"       "LUM"          "TUBA1B"       "S100A6"       "COL1A1"      
 [43] "LGALS1"       "COMP"         "MYH9"         "TYMP"         "MXRA8"        "ST14"        
 [49] "POSTN"        "RHOC"         "RBP1"         "MYL9"         "S100A4"       "MMP11"       
 [55] "CAPG"         "TNFRSF12A"    "GSTP1"        "FBN1"         "CD74"         "S100A13"     
 [61] "ACADVL"       "CTHRC1"       "FKBP10"       "COL1A2"       "ANXA2"        "C1QA"        
 [67] "VCAN"         "C1QB"         "NDUFS6"       "RHOB"         "FSTL1"        "MGAT4B"      
 [73] "TAX1BP3"      "THY1"         "RABAC1"       "C5orf46"      "COL12A1"      "CLTB"        
 [79] "YWHAH"        "ASPN"         "COL18A1"      "KRT7"         "HSPB1"        "TUBB2A"      
 [85] "CD151"        "ADAMTS2"      "MT-CO1"       "LOXL1"        "UQCR11"       "MT-CO3"      
 [91] "C1QC"         "CRIP2"        "AEBP1"        "CD99"         "TPM2"         "LMNA"        
 [97] "TUBB3"        "FLNA"         "S100A1"       "COL5A1"       "CRABP2"       "MMP7"        
[103] "TMEM176A"     "KRT23"        "NME2"         "S100A11"      "SELENOW"      "UBL5"        
[109] "S100A1"       "TMEM176B"     "LSR"          "S100A14"      "CLDN4"        "KRT8"        
[115] "TAGLN2"       "CLDN3"        "FRZB"         "CLU"          "CREG1"        "TMSB4X"      
[121] "S100A16"      "CALB2"        "PRDX5"        "PRELID1"      "IL34"         "S100A13"     
[127] "ADGRG1"       "DDA1"         "KRT7"         "PODXL"        "TMSB10"       "TMEM256"     
[133] "IFITM3"       "CLDN7"        "PYCARD"       "DEPP1"        "PRSS8"        "KLF6"        
[139] "TMED9"        "CRYAB"        "JUP"          "DMKN"         "COX6B1"       "FABP5"       
[145] "MGST3"        "CHI3L2"       "ALKBH7"       "SDC1"         "FXYD3"        "S100A6"      
[151] "TMEM219"      "ANXA2"        "LAMTOR2"      "MRPL36"       "PCSK1N"       "LMAN2"       
[157] "LGALS1"       "HIGD2A"       "NDUFV2"       "SRI"          "MYL12A"       "MYL12B"      
[163] "RBP1"         "RCSD1"        "KRT18"        "TPT1-AS1"     "KRT19"        "ATP5ME"      
[169] "SIAE"         "SUB1"         "MICOS13"      "RGS10"        "NDUFA1"       "GABARAP"     
[175] "OST4"         "NDUFA7"       "VTCN1"        "A1BG"         "NIPSNAP1"     "CCND1"       
[181] "RAB32"        "ARF5"         "HSPB1"        "GPI"          "TMEM147"      "MTLN"        
[187] "CAPG"         "DNAJC15"      "WDR83OS"      "ACTG2"        "S100A9"       "RPS27L"      
[193] "TSC22D1"      "TMEM258"      "TCEAL4"       "FAM83H"       "PSMB4"        "SEM1"        
[199] "SHANK2"       "CLTB"         "TMEM123"      "FARSA-AS1"    "CLU"          "CCDC71L"     
[205] "HIST1H1C"     "SOD2"         "ARF5"         "TMEM176A"     "NAMPT"        "SYPL1"       
[211] "RPL36"        "NUPR1"        "TRIM2"        "DRAP1"        "KRT8"         "PRELP"       
[217] "TOB1"         "SLC7A5"       "FZD8"         "RPL8"         "ZFP36L2"      "JUND"        
[223] "RHEB"         "FTH1"         "SMARCD3"      "FTL"          "ESYT2"        "MIF"         
[229] "MTHFD2"       "BHLHE40"      "SELENOM"      "CD9"          "KRT19"        "RPL28"       
[235] "TSTD1"        "MATN2"        "STC2"         "NME2"         "KHDRBS3"      "SPON2"       
[241] "CNN3"         "RPS19BP1"     "GPNMB"        "UAP1"         "BOD1"         "VEGFA"       
[247] "SLC39A4"      "LY6D"         "PRSS8"        "SBSN"         "RPL18"        "DNAJB1"      
[253] "CEBPD"        "SQSTM1"       "JUNB"         "COX6B1"       "SARS"         "TMEM176B"    
[259] "LBP"          "A1BG"         "EFNA1"        "CEBPG"        "GUK1"         "KRT16"       
[265] "AK4"          "PLEKHF1"      "CITED2"       "SLURP1"       "GFPT2"        "STOM"        
[271] "TNFRSF1A"     "SERPINH1"     "ADAMTS1"      "SOD3"         "STMP1"        "YIPF2"       
[277] "STEAP3"       "ALDOA"        "GPT"          "PYCR1"        "MYC"          "GFPT1"       
[283] "IFRD1"        "HILPDA"       "MEAF6"        "SPIRE1"       "HSP90AB1"     "SERPINA3"    
[289] "FRZB"         "FADS3"        "SUSD2"        "CLDN15"       "RPS8"         "AP000851.2"  
[295] "SEZ6L2"       "YY1"          "TM4SF1"       "CAPNS1"       "RPS28"        "WIPI1"       
[301] "RPS23"        "PABPC1"       "EEF1A1"       "RPL30"        "CRLF1"        "RPS6"        
[307] "RPL7A"        "RPS18"        "GAS5"         "AZGP1"        "FBXO32"       "RPS10"       
[313] "RPL32"        "SOX4"         "AARD"         "MIR205HG"     "EEF1D"        "RPL12"       
[319] "RPS15A"       "MTRNR2L12"    "RPS29"        "NCAN"         "TMSB15A"      "MALAT1"      
[325] "S100B"        "CPE"          "WIF1"         "RPL14"        "ROPN1"        "DEK"         
[331] "MT2A"         "RPL23A"       "IFITM1"       "ACTG1"        "RPL35A"       "RGS2"        
[337] "RPL15"        "COL2A1"       "FABP7"        "LTF"          "NRN1"         "NSA2"        
[343] "IGHG3"        "RPS3"         "MDK"          "S100P"        "IGKC"         "IL6"         
[349] "ASPH"         "SPOCK2"       "COL9A3"       "PTGES"        "RFLNB"        "PEG10"       
[355] "LDHB"         "EIF3H"        "NAP1L1"       "BTF3"         "TBCA"         "RPL21"       
[361] "RPS24"        "CD44"         "PACRG"        "RPL39"        "P4HA1"        "RPL37A"      
[367] "PLIN2"        "PTGFR"        "PRPH"         "RAD21"        "LEFTY2"       "CA9"         
[373] "TMEM139"      "SELENOP"      "RPL41"        "GLRX"         "ASPM"         "MTRNR2L8"    
[379] "WNT6"         "LHX3"         "RBP7"         "ANP32B"       "CP"           "MOXD1"       
[385] "NDRG1"        "RPL38"        "CTNND2"       "PNRC1"        "ITGA9"        "RHOBTB3"     
[391] "CCDC3"        "CCNE2"        "EPB41L4A-AS1" "TCF19"        "SNORC"        "AL583785.1"  
[397] "INTS13"       "INAVA"        "RPS3A"        "TUBB"  
```

```{r}
topgenes_heatmap <- comparePolygonExpression(visium_sample_119B, polygon_name = "selections_multi_regions",
                         selected_feats = top30_genes, method = "scran", expression_values = "normalized")
# This function do not save the plot
topgenes_heatmap
pdf("/projectnb/rd-spat/HOME/ivycwf/project_1/sample_119B/top30genes_heatmap_multi_regions.pdf",width = 8, height = 15)
plot(topgenes_heatmap)
dev.off()
```

```{r}
library(SeqGSEA)
#Convert symbol to ensembl ID
 adipo1_ensem <- SeqGSEA::convertSymbol2Ensembl(top_genes[1:100])%>%
   dplyr::select(ensembl_gene_id)%>%
   as.vector()
 
```

```{r}
#Convert symbol to ensembl ID
 adjacent_ensem <- SeqGSEA::convertSymbol2Ensembl(top_genes[101:200])%>%
   dplyr::select(ensembl_gene_id)%>%
   as.vector()

```

```{r}
#Convert symbol to ensembl ID
 far_ensem <- SeqGSEA::convertSymbol2Ensembl(top_genes[201:300])%>%
   dplyr::select(ensembl_gene_id)%>%
   as.vector()

```


```{r}
save.image(file = "/projectnb/rd-spat/HOME/ivycwf/project_1/sample_119B/visium_119B.RData")
```

```{r}
load(file ="/projectnb/rd-spat/HOME/ivycwf/project_1/sample_119B/visium_119B.RData" )
```







# ###############################################################################
# Plot with group by: (optional)
```{r}
spatPlot2D(visium_sample_95B, cell_color = 'leiden_clus',
           group_by = 'leiden_clus', coord_fix_ratio = 1,
           cow_n_col = 6, show_legend = F,
           save_param = list(base_width = 14, base_height = 14))
#Highlight one or more groups: (optional)
spatPlot2D(visium_sample_95B, cell_color = 'leiden_clus',
           select_cell_groups = '8', coord_fix_ratio = 1, show_other_cells = TRUE,
           cell_color_code = c('8' = 'red'), other_cell_color = "grey", other_point_size = 1.5,
           save_param = list(base_width = 7, base_height = 7))
```

# Part 5: subset data (Do not need right now )
```{r}
# create and show subset
DG_subset = subsetGiottoLocs(visium_sample_95B,
                             x_max = 6500, x_min = 3000,
                             y_max = -2500, y_min = -5500,
                             return_gobject = TRUE)
# (x and y means the coordinate on the plots)

spatDimPlot(gobject = DG_subset,
            cell_color = 'leiden_clus', spat_point_size = 5)
```


```{r}
full_image<-createGiottoLargeImage("/projectnb/rd-spat/HOME/ivycwf/project_1/resolution/119B.tif")

plot(full_image)




scale_json_path = '/projectnb/rd-spat/HOME/ivycwf/GSE210616/119B/GSM6433605_119B_scalefactors_json.json'
scalefactors = jsonlite::read_json(scale_json_path)



# Generates deep copy of SpatRaster
full_image<-createGiottoLargeImage("/projectnb/rd-spat/HOME/ivycwf/project_1/resolution/119B.tif")

#put an large image in the Giotto object
visium_sample_119B <- setGiottoImage(visium_sample_119B, image = "largeImage")

visium_sample_119B@largeImages <- list(image = full_image)

showGiottoImageNames(visium_sample_119B) #know the information of the images store in the Giotto object 

#Get all the spatial location
sl= getSpatialLocations(visium_sample_119B)

#Get a test cell have smallest x coordinate
testcell = sl[][sdimx == min(sdimx),]

#Get the scale factors that is used here
scale_json_path = '/projectnb/rd-spat/HOME/ivycwf/GSE210616/119B/GSM6433605_119B_scalefactors_json.json'

#Read in the scale factor file
scalefactors = jsonlite::read_json(scale_json_path)


fullsize_sr = full_image@raster_object

save.image(file = "/projectnb/rd-spat/HOME/ivycwf/project_1/sample_119B/visium_119B_full_image.RData")


#$fiducial_diameter_fullres -- 324.4803

extent_region <- testcell$sdimx+
patch_size <- extent(testcell)



